---
title: "Bike Share Data"
output: github_document
---

- [ ] Which roads have high bike traffic? (Jan 1 2020 - Dec 31 2020)
    - [x] Which Capital Bikeshare Station pairs have the highest traffic?
    - [ ] What is the most direct route between those stations? 
    - [ ] How long does the most direct take by bike (predicted)?
    - [ ] How long is the median most direct ride time (actuals)?
    - [ ] Based on the predicted direct ride time and median recorded ride time, how many bicyclists are traveling directly from Station A to Station B? 

# Packages and Data

```{r}

library(tidyverse)
library(osrmr)

load("data/bike_data.RData")

```

# Busiest station pairs

```{r}
stations_traffic <- bikerides %>%
  group_by(start_station_name,end_station_name) %>%
  summarise(n = n()) %>%
  filter(start_station_name != end_station_name) %>% #filter out roundtrips that begin and end at a single station
  arrange(desc(n)) %>%
  view()

# this data source may not be as useful as planned....
```

#Most direct routes


```{r}

```

