---
title: "Bike Incidents & Existing Lanes"
output: github_document
---

# Packages & Data

```{r message = FALSE}
library(tidyverse)
library(ggmap)
library(ggspatial)
library(tidyverse)
library(sf)
library(geosphere)
library(readxl)

load("data/bike_data.RData")
```

# Visualizaing

```{r message = FALSE}
map2 <- get_map(location = moco_bb)

ggmap(map2) +
  geom_point(data = bike_incidents, aes(x = Longitude, y = Latitude, color = `Injury Severity`)) +
  layer_spatial(osm_bike$osm_lines$geometry) + 
  labs(title = "Bike Incident Severity Map with Bike Lanes") +
  theme(legend.position = c(0.8,0.8))

```

# Find road lenghts

This makes the assumption that if a cycleway (bike lane or trail) exists, the cyclist was using it.

```{r}

# Length of road data (for incident per mile calculation)
streets <- as(osm_streets$osm_lines,"SpatialLinesDataFrame")

streets_df <- data.frame(streets@data)


#Each road has multiple segments on our list because its characteristics can change
#loop through each to find all the segment ids, then add up all the segment lengths

road_lengths <- NULL

for (i in matched_streets$osm_name) {

street <- streets_df %>%
  filter(toupper(name) == i) 

road_length <- 0

  for (j in street$osm_id)  {

  x <- paste0("streets@lines$`",j,"`@Lines")

  segment <- lengthLine(eval(parse(text = x))[[1]]@coords)

  road_length <- road_length + segment

  
  }
  road_name <- street$name
  road_df <- data.frame(road_name,road_length)
  road_lengths <- bind_rows(road_lengths,road_df)

}

#it worked but there's a row for each segment instead of a row for each street. Also need to convert to miles.

road_lengths <- road_lengths %>%
  distinct() %>%
  mutate(road_length = road_length * 0.00062137,
         road_name = toupper(road_name)) %>%
  rename(osm_name = road_name)

matched_streets <- matched_streets %>%
  inner_join(road_lengths) %>%
  mutate(incidents_per_mile = n / road_length,
         incidents_per_mile_annual = incidents_per_mile / 4)

#which streets have the highest accident rate?

accident_rate <- matched_streets %>%
  select(osm_name,moco_name,n,road_length,incidents_per_mile,incidents_per_mile_annual) %>%
  arrange(desc(incidents_per_mile_annual)) 

#lets filter out roads with 3 or fewer accidents just to focus on places that are higher impact

accident_rate_trimmed <- accident_rate %>%
  filter(n >= 3)

head(accident_rate_trimmed)

```


# EDA of joined incident and LTS data

Used QGIS and the NNJoin plugin to join the LTS data to the Crash Reporting dataset; each incident has been joined to the characteristics of the nearest road segment.

[LTS Methodology](https://montgomeryplanning.org/wp-content/uploads/2017/11/Appendix-D.pdf)

Revised LTS Key: 
    - LTS 0 – None (separated trails and breezeways)  
    - LTS 1 – Very Low (Neighborhood roads, suitable for children)  
    - LTS 2 – Low ("suitable for most adults")  
    - LTS 2.5 – Moderate Low  
    - LTS 3 – Moderate High (four lane road with bike lane)  
    - LTS 4 – High (>40mph road)  
    - LTS 5 – Very High ("few bicyclists will brave these roads")  
    - LTS 9 - (not defined in methodology document; appears to be freeways/ramps. Accidents reported on these roads are n=11, 9 of which are labeled as at crossings)  
    - TRANS - (not defined in methodology document; appears to be Transit station related?)  
    - P - (not defined in methodology document; appears to be parking lots)

```{r}

joined_lts_crash <- read_excel("data/joined_lts_crash.xlsx")

#The `distance` column is distance to nearest neighbor; let's remove those that are distant from any given street, allowing for GPS imprecision 

joined_lts_crash_filtered <- joined_lts_crash %>%
  filter(distance < 50) %>%
  filter(!is.na(join_LTS_REV)) %>%
  filter(join_LTS_REV != "TRANS") %>%
  filter(join_LTS_REV != "P")

order <- c("NO APPARENT INJURY",
           "POSSIBLE INJURY",
           "SUSPECTED MINOR INJURY",
           "SUSPECTED SERIOUS INJURY",
           "FATAL INJURY")

joined_lts_crash_filtered$`Injury Severity` <- factor(joined_lts_crash_filtered$`Injury Severity`,levels = order)
```

## LTS by incident severity

```{r}
ggplot(joined_lts_crash_filtered,aes(x=join_LTS_REV,fill=`Injury Severity`)) +
  geom_bar() +
  labs(title="Bike Incident Injury Severity by \n Level of Traffic Stress (LTS) of nearest street",
       x="LTS",
       y="Number of incidents",
       fill="Injury Severity") +
  theme(legend.title = element_blank(),
        legend.position = c(.875,.9),
        legend.background = element_rect(fill = NA, colour = NA))

```

## Injury Severity by other road characteristics

```{r warning = FALSE}
ggplot(joined_lts_crash_filtered,aes(x=join_AVG_SLOPE,y=`Injury Severity`)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.1) +
  labs(title = "Injury Severity by \n Average Slope of road segment",
       x = "Average Slope",
       y = "Injury Severity")

ggplot(joined_lts_crash_filtered,aes(x=as.double(join_SPEEDLIM),y=`Injury Severity`)) +
  geom_boxplot()  +
  geom_jitter(alpha = 0.1) +
  labs(title = "Injury Severity by \n speed limit of road segment",
       x = "Speed Limit",
       y = "Injury Severity")

ggplot(joined_lts_crash_filtered,aes(x=as.double(join_BKLNWIDTH),y=`Injury Severity`)) +
  geom_jitter(alpha=0.3,shape = 16) +
  labs(title = "Injury Severity by \n Bike Lane Width of road segment",
       x = "Bike Lane Width (in feet) \n (0 means no bike lane)",
       y = "Injury Severity")

```

# Association of Incident Severity and LTS

Since we don't have traffic data, we can't make assumptions about the impact of LTS on incident rate - we just don't have the denominator. We *can* make assumptions about incident severity. 

```{r}

set.seed(1)

#contingency table of LTS vs injury severity
table(joined_lts_crash_filtered$join_LTS_REV,joined_lts_crash_filtered$`Injury Severity`)

#distribution, row-wise
prop.table(table(joined_lts_crash_filtered$join_LTS_REV,joined_lts_crash_filtered$`Injury Severity`), margin = 1)

#Test for independence
#Chi-squared test
chisq.test(table(joined_lts_crash_filtered$join_LTS_REV,joined_lts_crash_filtered$`Injury Severity`))
#p = 0.1315, but too many values of <5 for the results to be valid

#Fishers test
#Uses monte carlo simulation when the contingency table is larger than 2x2 (simulate.p.value = T), which also means we can't get confidence intervals out of it.
fisher.test(joined_lts_crash_filtered$join_LTS_REV,joined_lts_crash_filtered$`Injury Severity`,simulate.p.value = T)
# p = 0.04348 
#With p-values of <0.05, we can reject the null hypothesis and assume that these LTS and Injury Severity are associated 
```

Next, test some other associations

```{r}
#weather - no association
fisher.test(joined_lts_crash_filtered$Weather,joined_lts_crash_filtered$`Injury Severity`,simulate.p.value = T)

#light status (daylight, dawn, etc) - highly associated (p = 0.005997) but n is very low
fisher.test(joined_lts_crash_filtered$Light,joined_lts_crash_filtered$`Injury Severity`,simulate.p.value = T)

#number of lanes - no association
fisher.test(joined_lts_crash_filtered$join_MP_LANES,joined_lts_crash_filtered$`Injury Severity`,simulate.p.value = T)

#speed limit - highly associated (p = 0.002999) - expected as speed limit is a factor in LTS
fisher.test(joined_lts_crash_filtered$join_SPEEDLIM,joined_lts_crash_filtered$`Injury Severity`,simulate.p.value = T)

#shared road - no association
fisher.test(joined_lts_crash_filtered$join_MP_SHARDRD,joined_lts_crash_filtered$`Injury Severity`,simulate.p.value = T)

#average slope - no association
fisher.test(joined_lts_crash_filtered$join_AVG_SLOPE,joined_lts_crash_filtered$`Injury Severity`,simulate.p.value = T)

# to get a confidence interval we need to reduce the result (injury) to two levels: injury or no injury
simplified_injuries <- joined_lts_crash_filtered %>%
  mutate(injury = case_when(
    `Injury Severity` == "NO APPARENT INJURY" ~ "No injury",
    `Injury Severity` == "POSSIBLE INJURY" ~ "Injury",
    `Injury Severity` == "SUSPECTED MINOR INJURY" ~ "Injury",
    `Injury Severity` == "SUSPECTED SERIOUS INJURY" ~ "Injury",
    `Injury Severity` == "FATAL INJURY" ~ "Injury")) %>%
  select(join_LTS_REV,injury) %>%
  table()

simplified_injuries

prop.test(simplified_injuries)
#this also uses chi-sq and still doesn't return a confidence interval

```

Try something different. Bring in accident rates on each street + avg (mean?) LTS of those streets? Would need to convert LTS to an int; remove P, TRANS, NA, etc.

(some manual cleaning done on xlsx, e.g. matching "1st St" to "FIRST ST")

```{r warning = FALSE}
LTS_streets <- read_excel("data/LTS.xlsx")

AVG_LTS_Street <- LTS_streets %>%
  mutate(STREET_NAME = paste(STREET_N_1,STREET_TYP)) %>%
  group_by(STREET_NAME) %>%
  mutate(AVG_LTS = mean(as.numeric(LTS_REV),na.omit=T)) %>%
  select(STREET_NAME,AVG_LTS) %>%
  distinct()

LTS_accident_rate_joined <- AVG_LTS_Street %>%
  full_join(accident_rate, by = c("STREET_NAME" = "moco_name")) %>%
  mutate(incidents_per_mile_annual = ifelse(!is.na(incidents_per_mile_annual),incidents_per_mile_annual,0)) #streets with no incidents have a rate of 0


#correlation of average LTS across a street and incident rate across a street. 
cor.test(LTS_accident_rate_joined$AVG_LTS,LTS_accident_rate_joined$incidents_per_mile_annual)
#this shows no correlation, so we can assume that at the broad "road" level there is no correlation between the average road LTS and the incident rate per mile; instead we can posit that accidents happen in clustered hotspots where the LTS is high, even if the average LTS along the whole of the road is lower.
#Additionally, the very large number of roads with zero accidents may mean that this test is not reliable/assumptions are not

LTS_accident_rate_joined %>%
  ggplot(aes(x=AVG_LTS,y=incidents_per_mile_annual)) +
  geom_point(alpha=0.5)


```


